Statistical performance analysis of neural mass models?

### 2013-06-30 manuscript/manuscript.tex
updated figures
added correct figures
edits to intro

@@

History of models and estimation attempts

@@

\paragraph{Neural Mass Model}
\begin{figure}[ht]
  \begin{center}
    \includegraphics{./figures/pdf/SingleMass.pdf}
  \end{center}
  \caption{\emph{\textbf{Neural Mass Model.} The somas of the neural mass are represented by the triangle. The synaptic connections are represented by the circle. The input mean firing rate, $g(v_m(t))$, is convolved with post-synaptic response kernel (PSRK), shown in the inset on the left, to give the mean membrane potential, $v_n(t)$. This is transformed by the sigmoid, shown in the inset on the right, to give the output mean firing rate, $g(v_n(t))$.}}
  \label{fig:SingleNeuralMass}
\end{figure}
Neural mass models map from a mean firing rate of a population of pre-synaptic neurons to a mean membrane potential of a post-synaptic population, which in turn determines that population's output firing rate. A graphical description of a neural mass model can be seen in Figure~\ref{fig:SingleNeuralMass}.

@@

The parameters of the model, $\tau$, $\alpha$, $v_0$ and $\varsigma$, can be set so the neural mass has characteristics of specific neural populations, such as pyramidal neurons, spiny stellate cells and fast and slow inhibitory interneurons (GABA$_\mathrm{a}$ and GABA$_\mathrm{b}$). The neural populations can then be connected to represent the circuitry of a cortical column and further to form networks of columns. Various kinds of neural mass models have been developed \cite{Silva1974,Jansen1995,Wendling2002,David2003}, which are depicted in Figure~\ref{fig:NMMs}. Each synaptic connection in these networks can be described by the 2$^{\mathrm{nd}}$-order system of Equation~\ref{eq:2ndOrderNMM}, where the resultant dimensions of the networks of neural mass models in Figures~\ref{fig:NMMs} a, b and c are 6, 12 and 18, respectively.
\begin{figure}[ht]
	\centering
		\includegraphics[scale=1]{./figures/pdf/NeuralMassesHoriz.pdf}
	\caption{\emph{\textbf{Models of Cortical Columns.} Each model has a pyramidal neural population, with various forms of feedback and inputs (Cx. = Cortex, Th. = Thalamus). The EEG is taken as the mean membrane potential of the pyramidal population, $v_p(t)$, with additive measurement noise.  a) Minimal model of a column with inhibitory feedback \cite{Silva1974}. b) An extension with excitatory feedback and afferent input from thalamus \cite{Jansen1995,David2003}. c) Column with inhibition occurring across two times scales, enabling modelling of higher frequency activity observed in seizures~\cite{Wendling2002}.}}
	\label{fig:NMMs}
\end{figure}

>>>>

\newcommand{\unit}[1]{\textcolor{green}{#1}}
\newcommand{\brian}[1]{\textcolor{blue}{#1}}

@@

Mesoscopic mathematical models of the brain are becoming increasingly popular as tools for generating hypotheses on the physical principals that govern neurodynamics. Forward models have led to theory on the emergence of the brains rhythms, the mechanisms of anesthetic agents, the brains response to sensory stimuli, short-term memory formation, visual hallucinations and epileptic seizures. Most recently, advances engineering methods and computing resources have expanded the utility of these models as integral elements of frameworks to estimate of system and parametric states that can not be directed measured using electrophysiology. An established framework to reliably estimate physical properties underlying brain dynamics has the potential to revolutionise how we observe, diagnose, treat and cure neurological disorders. For example, in epilepsy the dynamics of hidden system and parametric states are thought to be highly patient-specific and unknown. Thus the ability to estimate these quantities using the framework will enable a greater understanding of individuals’ pathologies, leading to more tailored therapies and better patient outcomes.  Furthermore, subject-specific models will enable the application of control theory to the neurosciences, which is typically limited to man-made systems.

To date a large body of work has demonstrated validity of these models in-principal, through the ability to reproduce experimental data. The next stage is understanding the performance of estimation algorithms. Once models have been fit to subjects a final step in assessing validity will be to examine the within subject predictive power.In order for such a framework to be established as a clinical platform further work is required for establishing validity and feasibility for the mesoscopic models. This paper introduces engineering tools to neuroscience community for formally accessing feasibility. We also postulate how these tools can be used to increase performance of estimation methods.

Influence on sampling rate and estimation?

 of this framework History of models and estimation attempts

@@

\paragraph{Neural Mass Model}
\begin{figure}[ht]
  \begin{center}
    \includegraphics{./figures/pdf/SingleMass_plos.pdf}
  \end{center}
  \caption{\emph{\textbf{Neural Mass Model.} The somas of the neural mass are represented by the triangle. The synaptic connections are represented by the circle. The input mean firing rate, $g(v_m(t))$, is convolved with post-synaptic response kernel (PSRK), shown in the inset on the left, to give the mean membrane potential, $v_n(t)$. This is transformed by the sigmoid, shown in the inset on the right, to give the output mean firing rate, $g(v_n(t))$.}}
  \label{fig:SingleNeuralMass}
\end{figure}
Neural mass models map from a mean firing rate of a population of pre-synaptic neurons to a mean membrane potential of a post-synaptic population, which in turn determines that population's output firing rate. A graphical description of a neural mass model can be seen in Figure~\ref{fig:SingleNeuralMass}.

@@

The parameters of the model, $\tau$, $\alpha$, $v_0$ and $\varsigma$, can be set so the neural mass has characteristics of specific neural populations, such as pyramidal neurons, spiny stellate cells and fast and slow inhibitory interneurons (GABA$_\mathrm{a}$ and GABA$_\mathrm{b}$). The neural populations can then be connected to represent the circuitry of a cortical column and further to form networks of columns. Various kinds of neural mass models have been developed \cite{Silva1974,Jansen1995,Wendling2002,David2003}, which are depicted in Figure~\ref{fig:NMMs}. Each synaptic connection in these networks can be described by the 2$^{\mathrm{nd}}$-order system of Equation~\ref{eq:2ndOrderNMM}, where the resultant dimensions of the networks of neural mass models in Figures~\ref{fig:NMMs} a, b and c are 6, 12 and 18, respectively.
\begin{figure}[ht]
	\centering
		\includegraphics[scale=1]{./figures/pdf/NeuralMassesHoriz_plos.pdf}
	\caption{\emph{\textbf{Models of Cortical Columns.} Each model has a pyramidal neural population, with various forms of feedback and inputs (Cx. = Cortex, Th. = Thalamus). The EEG is taken as the mean membrane potential of the pyramidal population, $v_p(t)$, with additive measurement noise.  a) Minimal model of a column with inhibitory feedback \cite{Silva1974}. b) An extension with excitatory feedback and afferent input from thalamus \cite{Jansen1995,David2003}. c) Column with inhibition occurring across two times scales, enabling modelling of higher frequency activity observed in seizures~\cite{Wendling2002}.}}
	\label{fig:NMMs}
\end{figure}
